---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header = TRUE)
```

And lets preview this data:

```{r}

head(inc)
```
```{r}

summary(inc)
```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

Some additional interesting information we can get from this data set is the number of unique values in certain columns.

For example, the number of unique values in the "Industry" column is:

```{r}
nrow(unique(inc["Industry"]))
```

And the number of unique values in the "State" column is:
```{r}
nrow(unique(inc["State"]))
```
Interestingly there are 52 unique values for the State column which tells us that most likely every state is being respresented and that there are two additional values that likely represent Washington DC and Puerto Rico.

Another interesting stat we can get from this data non-visually is checking the number of NA values in each column. This can be done by using the `colSums(is.na(inc))` function.

```{r}
colSums(is.na(inc))
```

Here we see that only the "Employees" column has NA values. This tells us we likely will not need to do much data cleaning for this data set.


## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}
# Answer Question 1 here
library(dplyr)
library(ggplot2)

# Answer Question 1 here
inc %>%
  count(State, n(), sort = T) %>%
  ggplot(aes(x = reorder(State, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(x = "State", y = "Number of Companies")
```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.


```{r}
incNy$EmployeesLog <- log(incNy$Employees)
# plot a violin plot
ggplot(incNy, aes(x = EmployeesLog, y = Industry, fill = Industry)) +
  geom_boxplot()
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}

# Answer Question 3 here
ggplot(inc, aes(x = Industry, y = log(Revenue / Employees), fill = Industry)) +
  geom_violin() +
  scale_y_continuous(name = "Revenue per Employee") +
  theme_minimal() +
  coord_flip()
```


